Thermodynamic fragments in magic state theories

### 2020-09-28 magic_paper.tex
Basic bounds on phi plus, phi minus

\author{Nikolaos Koukoulekidis}
	\affiliation{Department of Physics, Imperial College London, London SW7 2AZ, UK}
	\email{nk2314@imperial.ac.uk}
\author{David Jennings}
	\affiliation{School of Physics and Astronomy, University of Leeds, Leeds, LS2 9JT, UK}
	\affiliation{Department of Physics, Imperial College London, London SW7 2AZ, UK}

@@

The zoo of all magic operation classes is summarised in ~\cref{fig:zoo}.
Completely positive-Wigner-preserving operations~\cite{cit:wang} form the operation class $\Omax$.
Therefore, $\sigma$--fragments cover this theory of magic exactly and any magic subtheory is contained within this cover.
In particular, the stabilizer operations $\so$ are contained within $\Omax$.
\begin{figure}[t]
    \centering
        \includegraphics[scale=0.47
        ]{figs/operations.pdf}
    \caption{\textbf{Decomposition of a magic theory $\R$ intro $\sigma$--fragments.} 
    Examples of magic theories ($\so$: Stabilizer operations, $\Omax$: Completely positive-Wigner-preserving operations, $\rcu$: Random Clifford Unitaries -- subclass of $\so$) involve operations within the yellow regions, with all other established magic theories between these two.
    We introduce $\sigma$--fragments $\O_\sigma$ defined for all free states $\sigma$ that cover $\Omax$ with each one extendible to a set of stochastic maps outside the $\cptp$ operations.
    Within each $\sigma$--fragment $\bmd$-majorization can be used allowing for an \nick{in-depth approach} towards the study of magic state interconversion.
    }
    \label{fig:zoo}
\end{figure}

@@

\begin{theorem}\label{thm:sigmamajor}
    Let $\R = (\F, \O)$ be a theory of magic. Then $\rho \longrightarrow \tau$ is possible then  $\W{\tau} \prec_\sigma \W{\rho}$ within a $\sigma$--fragment.
    \nick{I understand that dropping the full-rank assumption on $\sigma$ makes the statement a lot stronger, but also incorrect e.g. in the case of the replacement channel $\E(\rho) = \ketbra{0}$, since $d$-majorization in $\O_{\ketbra{0}}$ is not well-defined.}
\end{theorem}
\begin{proof}
Suppose we can convert $\rho$ into $\tau$ in the magic theory. 
Thus there is some $\O_\sigma$ and some $\E \in \O_\sigma$ such that $\E(\rho) = \tau$, and $\E(\sigma)=\sigma$. 
Therefore, the Wigner distribution of this free operation satisfies $\W{\E} \in \stochw$ and $\W{\E}\W{\rho} = \W{\tau}$. 
Since $\sigma$ is full-rank and free, its Wigner distribution is strictly positive in all components, so it directly follows from~\cref{def:dmajor} that $\W{\tau} \prec_{\sigma} \W{\rho}$.
\end{proof}
This result can be understood as an extension of the idea of a magic monotone, where we replace $\M(\tau) \leq \M(\rho)$ with $\W{\tau} \prec_\sigma \W{\rho}$. 
However the majorization constraints can be used to place upper bounds on magic state distillation in a way that allows one to incorporate the physics of the allowed operations -- this enables us to bound how much magic can be distilled via quantum operations that, for example, preserve the equilibrium state of the system, or via operations that are symmetric about the $Z$-axis of the Bloch sphere.

@@

>>>>

\author{Nikolaos Koukoulekidis}
	\email{nk2314@imperial.ac.uk}
	\affiliation{Department of Physics, Imperial College London, London SW7 2AZ, UK}
\author{David Jennings}
	\affiliation{School of Physics and Astronomy, University of Leeds, Leeds, LS2 9JT, UK}
	\affiliation{Department of Physics, Imperial College London, London SW7 2AZ, UK}

@@

The zoo of all magic operation classes is summarised in ~\cref{fig:zoo}.
Completely positive-Wigner-preserving operations~\cite{cit:wang} form the operation class $\Omax$.
Therefore, $\sigma$--fragments cover this theory of magic exactly and any magic subtheory is contained within this cover.
In particular, the stabilizer operations $\so$ are contained within $\Omax$.
\begin{figure}[t]
    \centering
        \includegraphics[scale=0.47
        ]{figs/operations.pdf}
    \caption{\textbf{Decomposition of a magic theory $\R$ intro $\sigma$--fragments.} 
	Examples of magic theories ($\so$: Stabilizer operations, $\Omax$: Completely positive-Wigner-preserving operations, $\rcu$: Random Clifford Unitaries -- subclass of $\so$) involve operations denoted by the two yellow regions, with every other established magic theory  between them.
    We introduce $\sigma$--fragments $\O_\sigma$ defined for all free states $\sigma$ that cover $\Omax$. 
    Each $\O_\sigma$ is extensible to a set of stochastic maps outside the $\cptp$ operations.
    Within each $\sigma$--fragment $\bmd$-majorization can be used allowing for an intricate, but tractable approach towards the study of magic state interconversion.
    }
    \label{fig:zoo}
\end{figure}

@@

\begin{theorem}\label{thm:sigmamajor}
    Let $\R = (\F, \O)$ be a theory of magic. Then $\rho \longrightarrow \tau$ is possible iff  $\W{\tau} \prec_\sigma \W{\rho}$ within a $\sigma$--fragment.
    \nick{I understand that dropping the full-rank assumption on $\sigma$ makes the statement a lot stronger, but also incorrect e.g. in the case of the replacement channel $\E(\rho) = \ketbra{0}$, since $d$-majorization in $\O_{\ketbra{0}}$ is not well-defined.}
\end{theorem}
\begin{proof}
Suppose we can convert $\rho$ into $\tau$ in the magic theory. 
Thus there is some $\O_\sigma$ and some $\E \in \O_\sigma$ such that $\E(\rho) = \tau$, and $\E(\sigma)=\sigma$. 
Therefore, the Wigner distribution of this free operation satisfies $\W{\E} \in \stochw$ and $\W{\E}\W{\rho} = \W{\tau}$. 
Since $\sigma$ is full-rank and free, its Wigner distribution is strictly positive in all components, so it directly follows from~\cref{def:dmajor} that $\W{\tau} \prec_{\sigma} \W{\rho}$.
\end{proof}
This result can be understood as an extension of the idea of a magic monotone, where we replace $\M(\tau) \leq \M(\rho)$ with $\W{\tau} \prec_\sigma \W{\rho}$. 
However the majorization constraints can be used to place upper bounds on magic state distillation in a way that allows one to incorporate the physics of the allowed operations -- this enables us to bound how much magic can be distilled via quantum operations that, for example, preserve the equilibrium state of the system, or via operations that are symmetric about the $Z$-axis of the Bloch sphere.

@@

\subsection{Strange state Lorenz curves in the unital fragment}

In this section we analytically investigate the behaviour of the Lorenz curve of an $n$-copy Strange state in the unital fragment, 
\begin{equation}\label{eq:lcsu}
	\lc{\rho_{\rm{S}}(\epsilon)^{\otimes n}} {\frac{1}{d}\id}(x).
\end{equation}
An analytic handle on this type of Lorenz curves, which we call ``standard'' Lorenz curves, elevates Strange states to ``standard candles'' in the context of magic state distillation \nick{expand}.
We first present some relevant theory on error bounds of general binary codes and then show how it can be used to approximate the Lorenz curves of~\cref{eq:lcsu}.

\subsubsection{Binomial distributions and error bounds}
In this section, for any non-negative reals $a,b$ with $b\geq a$, we will denote the set of integers $n$ such that $a\leq n \leq b$ as $[a,b]$ for convenience.

Consider an experiment consisting of $n$ trials of throwing a $p$-coin, that is a coin with probability $p$ of landing on one side and $1-p$ of landing on the other.
That could be a string of $n$ bits with bit-flip error probability $p$.

The experiment is described by binomial statistics and we define the ``left'' and ``right'' tails of the cumulative binomial distribution,
\begin{align}
	\Phi_\ell(m; n, p) &\coloneqq \sum\limits_{j=0}^m \binom{n}{j} p^j (1-p)^{n-j}, \label{eq:phi}\\
	\Phi_r(m; n, p) &\coloneqq \sum\limits_{j=m}^n \binom{n}{j} p^j (1-p)^{n-j}, \label{eq:phir}
\end{align}
where $m\in [0,n]$.

They are related by the symmetry between the left tail of a $p$-coin distribution and the right tail of a $(1-p)$-coin distribution,
\begin{equation}\label{eq:phil2phir}
	\Phi_\ell(m; n, p) = \Phi_r(n-m; n, 1-p),\ m\in [0,n],
\end{equation}
as well as the normalisation condition,
\begin{equation}
	\Phi_\ell(m; n, p) + \Phi_r(m+1; n, p) = 1,\ m\in [0,n-1].
\end{equation}

Equivalently, we get
\begin{align}
	\Phi_\ell(m; n, p) + \Phi_\ell(n-m-1; n, 1-p) &= 1,\ m\in [0,n-1], \label{eq:reversephi}\\
	\Phi_r(m; n, p) + \Phi_r(n-m+1; n, 1-p) &= 1,\ m\in [1,n]. \label{eq:reversephir}
\end{align}

The classical entropy of a $p$-coin is
\begin{equation}\label{eq:ent}
	S(p) \coloneqq -p\log{p} -(1-p)\log{(1-p)},
\end{equation}
which is symmetric, $S(p) = S(1-p)$.
%Let $c(m; n) \coloneqq [m(1-m/n)]^{-1/2},\ m\in [1, n-1]$.
A useful result is the entropic bound on a combination~\cite{cit:ash}.
\begin{lemma}\label{lem:comb_bounds}
	For all $m\in [1,n-1]$,
	\begin{align}
		&\left[ 8m\left(1-\frac{m}{n}\right) \right]^{-\frac{1}{2}} 2^{nS\left(\frac{m}{n}\right)} \leq \binom{n}{m} \leq \\
		&\left[ 2\pi m\left(1-\frac{m}{n}\right) \right]^{-\frac{1}{2}} 2^{nS\left(\frac{m}{n}\right)}.
	\end{align}
\end{lemma}
\begin{proof}
	For $m = 1,2, n-1, n-2$ check by direct calculation.	
	For all other cases, use Stirling's approximation.
\end{proof}

Bounds on the tails of a cumulative binomial distribution can be expressed in terms of the classical relative entropy between a $p$-coin and a $q$-coin,
\begin{equation}\label{eq:ent_rel}
	\ent{p}{q} \coloneqq p \log{\frac{p}{q}} + (1-p) \log{\frac{1-p}{1-q}}.
\end{equation}
This entropy satisfies the identity
\begin{equation}
	\ent{p}{q} = \ent{1-p}{1-q}.
\end{equation}
We can now derive entropic bounds on binomial cumulative distributions~\cite{cit:ash}.
\begin{lemma}\label{lem:phir_bounds}
	Given fixed $n>0$ and $p$, $\Phi_r$ satisfies the following bounds:
	\begin{align*}
		\begin{split}
		&\text{1. } \Phi_r(m; n, p) \geq \left[ 8m\left(1-\frac{m}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ents},\ m\in [1,n-1] \\
		&\text{2. } \Phi_r(m; n, p) \leq 2^{-n\ents},\ m\in [np,n] \\
		&\text{3. } \Phi_r(m; n, p) \leq 1 - \left[ 8(m-1)\left(1-\frac{m-1}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{m-1}{n}}{p}}, \\
		&\hspace{14pt} m\in [2,n]
		\end{split}
		\\
		&\text{4. } \Phi_r(m; n, p) \geq 1 - 2^{-n\ent{\frac{m-1}{n}}{p}},\ m\in [2,np+1]
	\end{align*}

	The asymptotic behaviour of $\Phi_r$ is determined by the relative entropy, 
	\begin{equation}
		\lim_{n\rightarrow\infty}{-\frac{1}{n}\log{\Phi_r(m; n, p)} = \ent{\frac{m}{n}}{p}},\ m\in [np,n]
	\end{equation}
\end{lemma}
\begin{proof}
	For bound (1), note that
	\begin{align*}
		\Phi_r(m; n, p) &= \sum\limits_{j=m}^n \binom{n}{j} p^j (1-p)^{n-j} \\
						   &\geq \binom{n}{m} p^{m} (1-p)^{n-m} \\
						   &\geq \left[ 8m\left(1-\frac{m}{n}\right) \right]^{-\frac{1}{2}} 2^{nS\left(\frac{m}{n}\right)} p^{m} (1-p)^{n-m} \\
						   &= \left[ 8m\left(1-\frac{m}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ents}.
	\end{align*}
	The first inequality follows from the positivity of the terms in the series, the second from~\cref{lem:comb_bounds} and the equality from the entropy definitions in~\cref{eq:ent,eq:ent_rel}

	For bound (2), consider a binomial random variable $X = \sum_{i=1}^N$, where $X_i$ takes the values 1, 0 with probabilities $p$ and $1-p$ respectively.
	Then we have
	\begin{align*}
		\Phi_r(m; n, p) &= P(X \geq m) = P\left(2^{rX} \geq 2^{rm}\right) \\
						   &\leq 2^{-rm} E\left(2^{rX}\right) = 2^{-rm + n\log{(1-p+pe^r)}} \\
						   &\leq 2^{-n\ents}.
	\end{align*}
	The first inequality holds for any discrete random variable $X$ with $r \geq 0$, while the second inequality involves maximisation of the exponent which occurs at $e^r = \frac{m(1-p)}{(n-m)p}$ (a non-negative $r$ exists if $np \leq m$).

	Bound 3 follows from bound 1 and~\cref{eq:reversephir}. Bound 4 follows from bound x2 and~\cref{eq:reversephir}.

	The asymptotic limit is a direct consequence of the squeeze theorem.
\end{proof}

\begin{lemma}\label{lem:phil_bounds}
	Given fixed $n>0$ and $p$, $\Phi_\ell$ satisfies the following bounds:
	\begin{align*}
		\begin{split}
		&\text{1. } \Phi_\ell(m; n, p) \geq \left[ 8m\left(1-\frac{m}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ents},\ m\in [1,n-1] \\
		&\text{2. } \Phi_\ell(m; n, p) \leq 2^{-n\ents},\ m\in [0,np] \\
		&\text{3. } \Phi_\ell(m; n, p) \leq 1 - \left[ 8(m+1)\left(1-\frac{m+1}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{m+1}{n}}{p}}, \\
		&\hspace{14pt} m\in [0,n-2]
		\end{split}
		\\
		&\text{4. } \Phi_\ell(m; n, p) \geq 1 - 2^{-n\ent{\frac{m+1}{n}}{p}},\ m\in [np+1,n-2]
	\end{align*}

	The asymptotic behaviour of $\Phi_\ell$ is determined by the relative entropy,
	\begin{equation}
		\lim_{n\rightarrow\infty}{-\frac{1}{n}\log{\Phi_r(m; n, p)} = \ent{\frac{m}{n}}{p}},\ m\in [0,np]
	\end{equation}
\end{lemma}
\begin{proof}
	This lemma is a direct consequence of~\cref{lem:phir_bounds} and~\cref{eq:phil2phir}.
\end{proof}

We now define the alternating counterpart of the left tail distribution,
\begin{equation}\label{eq:phialt}
	\Phi_\ell^{\rm{alt}}(m; n, p) \coloneqq \sum\limits_{j=0}^m \binom{n}{j} (-p)^j (1-p)^{n-j},
\end{equation}
where $m\in [0,n]$.
$\Phi_\ell^{\rm{alt}}$ is \emph{not} the cumulative distribution of any random variable.
It is upper bounded by $\Phi_\ell$,
\begin{equation}\label{eq:alt_bound}
	\abs{\Phi_\ell^{\rm{alt}}(m; n, p)} \leq \Phi(m; n, p),\ m\in [0,n]
\end{equation}
with equality iff $m=0$.

We further define the two balanced combinations of $\Phi_\ell, \Phi_\ell^{\rm{alt}}$ for all $m\in [0,n]$,
\begin{align}	
	\Phi_+(m; n, p) &\coloneqq \frac{1}{2} \left( \Phi_\ell(m; n, p) + \Phi_\ell^{\rm{alt}}(m; n, p) \right), \\
	\Phi_-(m; n, p) &\coloneqq \frac{1}{2} \left( \Phi_\ell(m; n, p) - \Phi_\ell^{\rm{alt}}(m; n, p) \right), \\
	\Phi(m; n, p) &\coloneqq \label{eq:phi}
	\begin{cases}
		\Phi_+(m; n, p),\ \text{for even }m\in[0,n] \\
		\Phi_-(m; n, p),\ \text{for odd }m\in[0,n].
	\end{cases}
\end{align}
\begin{lemma}\label{lem:phi_bounds}
	Given fixed $n>0$ and $p$, $\Phi$ satisfies the following bounds:
	\begin{align*}
		\begin{split}
		&\text{1. } \Phi(m; n, p) \geq \left[ 8m\left(1-\frac{m}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ents},\ m\in [1,n-1] \\
		&\text{2. } \Phi(m; n, p) \leq 2^{-n\ents},\ m\in [0,np] \\
		&\text{3. } \Phi(m; n, p) \leq 1 - \left[ 8(m+1)\left(1-\frac{m+1}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{m+1}{n}}{p}}, \\
		&\hspace{14pt} m\in [0,n-2]
		\end{split}
		\\
		&\text{4. } \nick{!!!}
	\end{align*}
\end{lemma}
\begin{proof}
	Bound 1 follows directly from~\cref{lem:comb_bounds}.

	Bounds 2,3 follow from bounds 2,3 in~\cref{lem:phil_bounds} combined with~\cref{eq:phi,eq:alt_bound}.
\end{proof}

\subsubsection{Standard Lorenz curves in terms of binomial distributions}
The coordinates of the standard Lorenz curve can be directly expressed in terms of $\Phi_+, \Phi_-$, for all $m$, some fixed $p$ and number of copies $n$.

\null\newpage
\null\newpage
